# Organize Raw Files

Before starting, make sure the data is saved and organized as stated in the Overview and open RStudio by opening the *UseRGuide_DataCollection.Rproj* R Project file. 

**Scenario**: Data collection has finished for a two session study. We need to prepare and analyze data for a number of tasks, one of which is the Flanker task. 
We only want to analyze the data for subjects that completed ALL sessions. This first requires some organization of the raw data files. What we want to do is:

1) Move the raw data files from **Tasks/Session 1/1. Flanker/data** to **Data Files/subj/Flanker**.

2) Separate the raw files for subjects that did not complete ALL sessions from those that did. The **subjlist.csv** file contains a list of subjects that completed ALL sessions.

These two steps can easily be done with the following line of code. Just type this into the console

```{r eval = TRUE}

```

Notice how at first there are no files in **Data Files/subj/Flanker**. Then after executing this line of code, it is now populated with *.edat* files.

This function will perform the two steps for every task in the **Session 1** folder. To perform them for any other sessions just change it from **Session 1** to any **Session #**.

You will also see a folder called **Did not finish all sessions**. The **.edat** files for subjects that did not finish all sessions will be in here. 

If you are doing this step before *Data Collection* has finished, and you do not want to delete the files from the **Tasks** directory or separate out subjects then leave out the `remove = TRUE` argument (or set it to FALSE) and leave out the `subj.list = ` argument.


********

```{r echo=FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
```

********

**The next chapter is more optional but you will probably find it useful**
