# Complex Span Tasks

<link rel="stylesheet" href="https://use.fontawesome.com/releases/v5.8.1/css/all.css" integrity="sha384-50oBUHEmvpQ+1lW4y57PTFmhCaXp0ML5d60M1M7uH2+nqUivzIebhndOJK28anvf" crossorigin="anonymous">

```{r include=FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
```

Once you are ready to start analyzing data from the complex span tasks, you will first need to create a merged (**.emrg**) data file of all the individual **.edat** files E-Prime produces. You then will need to export that **.emrg** file as either a tab-delimited .txt file or a comma delimited .csv file. You can then import the .txt or .csv merged file into R. `readr::read_delim()` or `readr::read_csv()`

Once you have the data file imported into R as a dataframe object, you can then use the functions to easily create clean and tidy raw data files for the complex span tasks. Depending on your sample size, these functions may actually take a little while to execute.

----

## Operation Span

`raw_ospan()` will create a clean and tidy raw data file for the operation span task. The only arguments you need to specify are:

* __x__: The dataframe object

* __blocks__: How many blocks were administered? (1-3)

## Symmetry Span

`raw_symspan()` will create a clean and tidy raw data file for the symmetry span task. The only arguments you need to specify are:

* __x__: The dataframe object

* __blocks__: How many blocks were administered? (1-3)

## Rotation Span

`raw_rotspan()` will create a clean and tidy raw data file for the rotation span task. The only arguments you need to specify are:

* __x__: The dataframe object

* __blocks__: How many blocks were administered? (1-3)

----

## Example: Symmetry Span

This is an example R Script to create a tidy raw data file for the Symmetry Span task:

```{r eval = FALSE}
## Set up ####
## Load packages
library(readr)
library(dplyr)
library(here)
library(datawrangling)
library(englelab)

## Set import/output directories
import.dir <- "Data Files/Merged"
output.dir <- "Data Files"

## Set import/output files
task <- "SymSpan"
import.file <- paste(task, ".txt", sep = "")
output.file <- paste(task, "_raw.csv", sep = "")
##############

## Import Data
data_import <- read_delim(here(import.dir, import.file),
                          "\t", escape_double = FALSE, trim_ws = TRUE)

## Clean up raw data and save
data_raw <- raw_symspan(data_import, blocks = 2)

## Output Data
write_csv(data_raw, path = here(output.dir, output.file))

rm(list=ls())

```


## Tidy File

The resulting dataframe will contain both raw trial-level data and scored data. The columns in the data are as follows

**Raw Trial-Level Columns**

- **Subject**:  Subject ID column

- **Block**:  Which block? (1-3)

- **Trial**:  Within a block, what trial number. 

    Trial number refers to an entire set-size sequence (presentation of processing items, memory items, and recall screen)
    
- **SetSize**:  For the trial, what is the set-size?

- **SubTrial**:  Within a trial, there are processing items presented sequentially. 

    Sub-Trial refers to this sequential presentation. It also represents the order of responses on the recall screen

- **SubTrialProc**:   "Processing" or "Recall" portion of the task?

- **RT**:  Reaction time

- **Accuracy**:  Accuracy (provided for both "Processing" and "Recall")

- **Response**:  The subject's response (provided for both "Processing" and "Recall")

- **CorrectResponse**:  The correct response (provided for both "Processing" and "Recall")

**Task Scores**

- **Processing.total**:  The total number of processing items correctly answered

- **Recall.total**:  The total number of recall items correctly answered

- **[Task].Absolute**: The task score using the Absolute Scoring method

    i.e. SymSpan.Absolute or OSpan.Absolute. 

- **[Task].Partial**:  The task score using the Partial Scoring method

- **[Task].Partial_Block1**:  The score on block 1 using the Partial Scoring method

- **[Task].Partial_Block2**:  The score on block 2 using the Partial Scoring method

- **[Task].Partial_Block3**:  The score on block 3 using the Partial Scoring method

- **[Task].[Processing]ACC**: Proportion of processing task items correctly answered

    i.e. SymSpan.SymmetryACC or OSpan.MathACC. 

- **[Task].Avg[Processing]Time**: The average time to complete a processing task item

     i.e. SymSpan.AvgSymmetryTime or OSpan.AvgMathTime.
     


********

```{r echo=FALSE, message=FALSE, warning=FALSE, eval = TRUE}
rm(list=ls())
```

********

**The next Chapter is a summary of this Section**
