# Example Data: Flanker {-}

<link rel="stylesheet" href="https://use.fontawesome.com/releases/v5.8.1/css/all.css" integrity="sha384-50oBUHEmvpQ+1lW4y57PTFmhCaXp0ML5d60M1M7uH2+nqUivzIebhndOJK28anvf" crossorigin="anonymous">

```{r include=FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE, message = FALSE, warning = FALSE)
```

If you have not done so already, open RStudio by opening the **UseRGuide_DataAnalysis.Rproj** file. 

In the **Files Pane** navigate to **Data Files/Raw Data** and click on the **Flanker_raw.csv** file -> Import Data Set... -> Update (top right) -> Import.

Note: This is just a quick way to view a csv file. It is not suggested you import files this way.

Let's get to know this data set. Looks like there is one row per trial per subject. Both practice and real trials are included. There are *neutral*, *congruent*, and *incongruent* trials. Both RT and Accuracy are recorded. 

All this can easily be seen just by viewing the data file. Sometimes you cannot see all the information you want very easily just by viewing the file. This is where some basic R functions come in handy.

To evaluate the different TrialProcs:

```{r}
unique(Flanker_raw$TrialProc)
```

To evaluate how many different conditions there are:

```{r}
unique(Flanker_raw$Condition)
```

How many trials are there?

```{r}
length(unique(Flanker_raw$Trial))
```

How many subjects?

```{r}
length(unique(Flanker_raw$Subject))
```

What are all the column names?

```{r}
colnames(Flanker_raw)
```

I often times print `colnames()` in the console as I type code so that I can easily see the column names I need to refer to. 

It is a good idea to get familiar with a data set before working on it. By doing this you may even catch errors, typos, or other mistakes in the data.

********

```{r echo=FALSE, message=FALSE, warning=FALSE, eval = TRUE}
rm(list=ls())
```

********

**Woo! You should feel accomplished. Okay one more quick step before the best part.**
