# Plotting Means

```{r include=FALSE, eval = TRUE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, message = FALSE, warning = FALSE)
```

You can go ahead and set a default theme for your plots

```{r}
theme_set(theme_linedraw())
```


## Bar Graphs

Bar graphs are the standard. They are ubiquitous across psychology. Basically everyone uses them. But in all honesty, **Bar graphs SUCK!**. 

The worst part about them is that they hide the distribution of the raw data points (even when error bars are included). Even worse, too often you will see bar graphs with NO ERROR BARS! Yikes! A bar graph with no error bars tells you almost NOTHING!

To illustrate this let's use a data set containing information on mammalian sleep patterns from the data set `msleep`.

```{r}
head(msleep)
```

Let's plot the relationship between the different eating habits (`vore`) and total sleep time (`sleep_total`).


```{r}
msleep1 <- filter(msleep, !is.na(vore))

ggplot(msleep1, aes(vore, sleep_total)) +
  stat_summary(fun.y = mean, na.rm = TRUE, geom = "bar")
```

This only tells us what the means are. We have no idea about the distributions. Well for this reason people usually like to see error bars. Okay well let's add error bars.

```{r}
ggplot(msleep1, aes(vore, sleep_total)) +
  stat_summary(fun.y = mean, na.rm = TRUE, geom = "bar") +
  stat_summary(fun.data = mean_cl_normal, na.rm =TRUE, 
               geom = "errorbar", width = .2)
```

Okay better. But we still cannot see the underlying distribution.

Here is a crazy idea. What if we plotted the raw data points. Like we do with scatterplots! Whoa! What a concept

```{r}
ggplot(msleep1, aes(vore, sleep_total)) +
  geom_point()
```

When plotting raw data points with categorical variables on the x-axis it makes more sense to `jitter` the points so they are not all just lying on top of each other. 

```{r}
ggplot(msleep1, aes(vore, sleep_total)) +
  geom_point(position = position_jitter(width = .2))
```

Wow! Does this give you a completely different picture than the bar graph with error bars? It does to me! Especially look at the `insecti` and `omni` eating habits. There is definitely a bi-modal distribution happening there. 

From the bar graph with error bars, we might be fooled into thinking that the distributions for `carni` and `omnie` are pretty similar. But are they? Not at all!

**THIS IS WHY YOU SHOULD ALWAYS PLOT THE RAW DATA POINTS**

But means and error bars are also useful information so let's add those

```{r}
ggplot(msleep1, aes(vore, sleep_total)) +
  geom_point(position = position_jitter(width = .2)) +
  stat_summary(fun.y = mean, na.rm = TRUE, 
               geom = "point", color = "dodgerblue") +
  stat_summary(fun.data = mean_cl_normal, na.rm =TRUE, 
               geom = "errorbar", width = .2, color = "dodgerblue")
```

Another aesthetic option that is useful when we are plotting means and error bars ontop of raw data is the `alpha` aesthetic. This can allow us to make the raw data points more transparent, fade into the background a little more.

```{r}
ggplot(msleep1, aes(vore, sleep_total)) +
  geom_point(position = position_jitter(width = .2), alpha = .3) +
  stat_summary(fun.y = mean, na.rm = TRUE, 
               geom = "point", color = "dodgerblue", 
               size = 4, shape = "diamond") +
  stat_summary(fun.data = mean_cl_normal, na.rm =TRUE, 
               geom = "errorbar", width = .2, color = "dodgerblue")
```

```{r}
ggplot(msleep1, aes(vore, sleep_total)) +
  geom_point(position = position_jitter(width = .2), alpha = .3) +
  stat_summary(fun.y = mean, na.rm = TRUE,
               geom = "point", color = "dodgerblue", 
               size = 4, shape = "diamond") +
  stat_summary(fun.data = mean_cl_normal, na.rm =TRUE, 
               geom = "errorbar", width = .2, color = "dodgerblue") +
  stat_summary(fun.y = mean, na.rm = TRUE, aes(group = 1),
               geom = "line", color = "dodgerblue", 
               size = .75, shape = "diamond") 
```

### Multidimensional Bar Plot

```{r}
library(tidyr)
iris.long <- iris %>%
  mutate(Flower = row_number()) %>%
  gather("Part", "Inches", -Flower, -Species) %>%
  separate(Part, into = c("Part", "Measurement")) %>%
  arrange(Flower, Species) %>%
  select(Flower, Species, Part, Measurement, Inches)

head(iris.long)

ggplot(iris.long, aes(Measurement, Inches, group = Species, color = Species)) +
  geom_point(position = position_jitterdodge(jitter.width = .2, 
                                             dodge.width = .7), 
             alpha = .1) +
  stat_summary(fun.y = mean, na.rm = TRUE, 
               geom = "point", shape = "diamond",
               size = 4, color = "black", 
               position = position_dodge(width = .7)) +
  stat_summary(fun.data = mean_cl_normal, na.rm = TRUE, 
               geom = "errorbar", width = .2, color = "black",
               position = position_dodge(width = .7)) +
  scale_color_brewer(palette = "Set1")
```

********

```{r echo=FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
```

********

**Now let's move on to univariate plots**

