# Import and Output Data Files

<link rel="stylesheet" href="https://use.fontawesome.com/releases/v5.8.1/css/all.css" integrity="sha384-50oBUHEmvpQ+1lW4y57PTFmhCaXp0ML5d60M1M7uH2+nqUivzIebhndOJK28anvf" crossorigin="anonymous">

The previous chapter was more of a high-level overview of the organizational structure and workflow of preparing and analyzing data in R.

In the next few chapters we will dive into more of the details for actually creating R scripts at each of the stages of data preparation and analysis.

First let's talk about importing and outputting data files in R.

## RStudio Projects and here

You need to be using RStudio Projects for anything you do in R.

RStudio Projects allow you to open isolated instances of R and RStudio for each of your projects. In combination with the `here` package it will also provide a simple and fool proof way of specifying file paths.

For instance, I have an .Rproj file saved in my "UseR_Guide" folder. When I use `here()` it will output a file path to that location.

```{r message = FALSE, warning = TRUE}
library(here)
here()
```

You can then use a relative file path inside of `here()`.

```{r}
here("Data Files/Raw Data/flanker_raw.csv")
```

This is equivalent to

```{r}
here("Data Files", "Raw Data", "flanker_raw.csv")
```

<br>

I typically like to set the first argument as the relative file path and the second argument as the file name. This visually separates the file path and the file name, making your script easier to read.

```{r}
here("Data Files/Raw Data", "flanker_raw.csv")
```

You can then use `here()` directly in import and output functions:

```{r eval = FALSE}
import <- read_csv(here("Data Files/Raw Data", "flanker_raw.csv"))
```


```{r eval = FALSE}
write_csv(data, here("Data Files/Scored Data", "flanker_scored.csv"))
```

Every time you use `here()` you know that the file path will start at where you have your .Rproj file saved. Instead of messing around with working directories with `setwd()` or `getwd()`, just use `here()` and RStudio Projects. This becomes especially helpful when working with RMarkdown documents.

```{r echo=FALSE, message=FALSE, warning=FALSE}
rm(list=ls())
```


