# A Mainscript

<link rel="stylesheet" href="https://use.fontawesome.com/releases/v5.8.1/css/all.css" integrity="sha384-50oBUHEmvpQ+1lW4y57PTFmhCaXp0ML5d60M1M7uH2+nqUivzIebhndOJK28anvf" crossorigin="anonymous">

When you have a lot of R scripts it can become tedious to open up and source each script one at a time. Instead, what you can do is have a **mainscript** file that executes (or *sources*) each R script file from one place.

The mainscript file essentially just contains lines of code using the `source()` function. For instance a mainscript in a **Data Collection** repository might look like:

```{r eval=FALSE}
## fluid intelligence
source("R Scripts/fluid intelligence/rapm_raw.R")
source("R Scripts/fluid intelligence/lettersets_raw.R")
source("R Scripts/fluid intelligence/numberseries_raw.R")

## working memory capacity
source("R Scripts/working memory capacity/symspan_raw.R")
source("R Scripts/working memory capacity/rotspan_raw.R")
source("R Scripts/working memory capacity/runletter_raw.R")
source("R Scripts/working memory capacity/rundigit_raw.R")
source("R Scripts/working memory capacity/mentalcounters_raw.R")

## attention control
source("R Scripts/attention control/antisaccade_raw.R")
source("R Scripts/attention control/sact_raw.R")
source("R Scripts/attention control/visual_flankerDL_raw.R")
source("R Scripts/attention control/visual_stroopDL_raw.R")

## other
source("R Scripts/demographics_raw.R")
```

This script will *source* all these R scripts without having to open them.

For **Data Analysis** a mainscript can also help to organize your data analysis procedure and the order in which certain R scripts need to be ran. For instance:

```{r eval = FALSE}
#################################################
#------ 1. Scored data ------# 
#################################################

source("R Scripts/1_baseline_score.R")
source("R Scripts/1_gf_score.R")
source("R Scripts/1_wmc_score.R")
source("R Scripts/1_antisaccade_score.R")
source("R Scripts/1_sact_score.R")
source("R Scripts/1_visualarrays_score.R")
source("R Scripts/1_questionnaires_score.R")

rm(list = ls())
#############################################################
#------ 2. Create Final Merged Data File for Analysis ------# 
#############################################################

source("R Scripts/2_merge.R", echo = TRUE)

rm(list = ls())
################################
#------ 3. Data Analysis ------# 
################################

rmarkdown::render("R Scripts/3_MainAnalyses.Rmd", 
                  output_dir = "Results", 
                  output_file = "MainAnalyses.html")
rmarkdown::render("R Scripts/3_Reliabilities.Rmd", 
                  output_dir = "Results", 
                  output_file = "Reliabilities.html")
rmarkdown::render("R Scripts/3_Demographics.Rmd", 
                  output_dir = "Results", 
                  output_file = "Demographics.html")

rm(list = ls())
#################################################
```


----

